---
title: "Portfolio Fall 2021 Project data cleaning"
author: "Adam Paul"
date: "3/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> The purpose of this portfolio is to have one final attempt at data cleaning within R. Specifically, I'll work with the data from my project we ran in Fall 2021. The following passage is adapted from my thesis prospectus, which was written about this project.

The purpose of the study was to determine whether disclosure causally benefits students’ feelings of class and institutional belonging, their motivation, and their academic performance. To do this, I conducted a laboratory study designed to mimic an online course. Small groups comprised of 2-4 students from the same Introductory Psychology class were brought into the lab together to engage in a disclosure task with one another. 

In the experimental condition, I used an abbreviated version of the Fast Friends protocol (Aron et al., 1997), which employs questions of increasing intimacy. This was contrasted with a control condition which used similar procedures but with “small talk” questions that are not expected to yield quality, intimate disclosure. To mimic the experience of an online class, conversations occurred virtually via Zoom.  Afterward, I measured students’ belonging, motivation, and performance. Finally, I had students take a brief quiz on research methods.

I had three hypotheses:

H1: Students in the disclosure condition will report greater belonging, at both the institutional and classroom levels, than students in the small talk condition.

H2: Students in the disclosure condition will report greater subject area motivation than students in the small talk condition.

H3: Students in the disclosure condition will have higher scores on the quiz than students in the small talk condition.

## Set-up


```{r loading packages, echo=TRUE, message=FALSE}
library('reshape2')
library('dplyr')
library('Rcpp')
library('readr')
library('tidyverse')
library('psych')
```

##### Bringing in the data

This data was collected as a part of three different Qualtrics surveys, one that followed the manipulation, one that measured belonging, and a final one that contained the quiz.

```{r bringing in survey 1 data, echo=TRUE, collapse=TRUE, message=FALSE}
survey1 <- read_csv("data/Fall 2021 Survey 1.csv")

# Ensuring the dataframe is a tibble.
as_tibble(survey1)

#Taking a look at the table.
View(survey1)
```

```{r bringing in survey 2 data, echo=TRUE, collapse=TRUE, message=FALSE}
survey2 <- read_csv("data/Fall 2021 Survey 2.csv")

# Ensuring the dataframe is a tibble.
as_tibble(survey2)

#Taking a look at the table.
View(survey2)
```

```{r bringing in quiz data, echo=TRUE, collapse=TRUE, message=FALSE}
quiz <- read_csv("data/Fall 2021 Quiz.csv")

# Ensuring the dataframe is a tibble.
as_tibble(quiz)

#Taking a look at the table.
View(quiz)
```


Some cleaning was done in excel prior to ensure that participant ids are identical across surveys (e.g., 103C and 103 C are made to be identical)


As with Study 1 and 2, the data has already had its identifiers removed, but if it didn't, here is where that would be done. 

survey <- subset(survey, select= -c(StartDate, EndDate, Status,	IPAddress, Progress,
                Finished, RecordedDate, ResponseId, Durationinseconds, RecipientLastName, RecipientFirstName,
                RecipientEmail, ExternalReference, LocationLatitude, LocationLongitude,
                DistributionChannel, UserLanguage))

### Data merge

```{r survey 1 & 2 merge, echo=TRUE, collapse=TRUE, message=FALSE}
study1 <- survey1 %>% 
  full_join(survey2, by = "participant_id")

#Always make sure it's a tibble
as_tibble(study1)

View(study1)
```

```{r quiz data merging, echo=TRUE, collapse=TRUE, message=FALSE}
study1 <- study1 %>% 
  full_join(quiz, by = "participant_id")

#Always make sure it's a tibble
as_tibble(study1)

View(study1)
```

### Data cleaning!

Now that we have the data together, we can start cleaning it up. A big thing I learned from this is that consistency in naming makes everything easier (i.e., using study1 in my code consistently, and keeping consistent names like demog_XXX for demographics).


##### Gender dummy coding

First, I need to look at whether we have non-binary individuals in the dataset.

Creating gender dummy codes that I can read easily helps down the line, similar to what I did in Study 2.

Notably, I hadn't changed the variable names to the standardized demog_XXX that I use later. I used demograph_XXX which is longer than needed. Fortunately, we were using the same numbering.

```{r renaming gender variables, collapse=TRUE}
#male
study1 <- rename(study1, "demog_gender_male" = demograph_gender_2)

#female
study1 <- rename(study1, "demog_gender_female" = demograph_gender_1)

#trans
study1 <- rename(study1, "demog_gender_trans" = demograph_gender_31)

#twospirit
study1 <- rename(study1, "demog_gender_twospirit" = demograph_gender_32)

#genderqueer
study1 <- rename(study1, "demog_gender_genderqueer" = demograph_gender_33)

#other
study1 <- rename(study1, "demog_gender_other" = demograph_gender_34)

#no answer
study1 <- rename(study1, "demog_gender_no_answer" = demograph_gender_4)
```


```{r checking gender demographics}
study1 %>%
        filter(is.na(demog_gender_female), is.na(demog_gender_male)) %>%
        count(demog_gender_trans, demog_gender_twospirit, demog_gender_genderqueer, demog_gender_other, demog_gender_no_answer)

study1 %>%
count(demog_gender_trans, demog_gender_twospirit, demog_gender_genderqueer, demog_gender_other, demog_gender_no_answer)

```

We have one participant who is missing responses for our variables of interest in survey 2, so they lack a gender score. Additionally, there is one person who only answered trans.

Once again, we originally included this person into the dichotomous gender code, since we can't make any assumptions about them and would lose degrees of freedom. This may not be the best way to handle it, but it is how we did it before.

```{r Creating binary gender dummy code}
study1 <- study1 %>%
        mutate(
        gender_male_dummy=case_when(
                demog_gender_male == 1 ~ 1,
                demog_gender_female == 1 ~ 0
                ))
```

##### Race dummy coding

Sadly, I once again have to correct my original code for demog vs demograph. I much prefer the former, but it's a good task for looking over code and adapting it to what is essentially someone else's (a much less code savvy me) code.

To create the race dummy codes, I first need to create a multiracial composite so that I can avoid adding multiracial people to individual dummy codes. 

Because the individual race codes have NA in all unfilled spots, I need to replace them with 0's to get the code to work correctly.

```{r adding 0 to original code}

study1 <- study1 %>%
        mutate(demograph_race_51=case_when(
                demograph_race_51 == 1 ~ 1,
                is.na(demograph_race_51) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_52=case_when(
                demograph_race_52 == 1 ~ 1,
                is.na(demograph_race_52) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_59=case_when(
                demograph_race_59 == 1 ~ 1,
                is.na(demograph_race_59) ~ 0
                                       ))
study1 <- study1 %>%
        mutate(demograph_race_21=case_when(
                demograph_race_21 == 1 ~ 1,
                is.na(demograph_race_21) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_22=case_when(
                demograph_race_22 == 1 ~ 1,
                is.na(demograph_race_22) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_23=case_when(
                demograph_race_23 == 1 ~ 1,
                is.na(demograph_race_23) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_29=case_when(
                demograph_race_29 == 1 ~ 1,
                is.na(demograph_race_29) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_31=case_when(
                demograph_race_31 == 1 ~ 1,
                is.na(demograph_race_31) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_32=case_when(
                demograph_race_32 == 1 ~ 1,
                is.na(demograph_race_32) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_33=case_when(
                demograph_race_33 == 1 ~ 1,
                is.na(demograph_race_33) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_39=case_when(
                demograph_race_39 == 1 ~ 1,
                is.na(demograph_race_39) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_11=case_when(
                demograph_race_11 == 1 ~ 1,
                is.na(demograph_race_11) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_12=case_when(
                demograph_race_12 == 1 ~ 1,
                is.na(demograph_race_12) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_13=case_when(
                demograph_race_13 == 1 ~ 1,
                is.na(demograph_race_13) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_19=case_when(
                demograph_race_19 == 1 ~ 1,
                is.na(demograph_race_19) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_41=case_when(
                demograph_race_41 == 1 ~ 1,
                is.na(demograph_race_41) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_42=case_when(
                demograph_race_42 == 1 ~ 1,
                is.na(demograph_race_42) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_89=case_when(
                demograph_race_89 == 1 ~ 1,
                is.na(demograph_race_89) ~ 0
                                       ))

study1 <- study1 %>%
        mutate(demograph_race_99=case_when(
                demograph_race_99 == 1 ~ 1,
                is.na(demograph_race_99) ~ 0
                                       ))
```


With that task finished, we can work on making the multiracial composite.


```{r creating a multiracial sum}
study1 <- study1 %>%
        rowwise() %>%
        mutate(
                multiracial_sum=sum(c_across(demograph_race_41:demograph_race_89
        )))

```

```{r Creating multiracial dummy code}
study1 <- study1 %>%
        mutate(race_multiracial=case_when(
                multiracial_sum > 1 ~ 1,
                multiracial_sum == 1 ~ 0,
                multiracial_sum == 0 ~ 0
        ))

```

Now I can make the composites! I used the find and replace function here for the first time.



##### White composite

Composed of individuals who identified as European/European American, Middle Eastern/Middle Eastern American, Other White; or demograph_race_51, demograph_race_52, and demograph_race_59.

```{r White composite dummy code}
study1 <- study1 %>%
        mutate(
        race_white=case_when(
                demograph_race_51 | demograph_race_52 | demograph_race_59 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_white=case_when(
          race_white == 1 ~ 1,
          race_white == 0 ~ 0,
          is.na(race_white) ~ 0 
          ))

```

##### Black composite

Composed of individuals who identified as African American/Black, African, Caribbean, Other Black; or demograph_race_21, demograph_race_22, demograph_race_23, and demograph_race_29

```{r Black composite dummy code}
study1 <- study1 %>%
        mutate(
        race_black=case_when(
                demograph_race_21 | demograph_race_22 | demograph_race_23 | demograph_race_29 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_black=case_when(
          race_black == 1 ~ 1,
          race_black == 0 ~ 0,
          is.na(race_black) ~ 0 
          ))

```

##### Hispanic composite

Composed of individuals who identified as Mexican American/Chicano, Puerto Rican, Central American, Other Hispanic or Latino; or demograph_race_31, demograph_race_32, demograph_race_33, and demograph_race_39.

```{r Hispanic composite dummy code}
study1 <- study1 %>%
        mutate(
        race_hispanic=case_when(
                demograph_race_31 | demograph_race_32 | demograph_race_33 | demograph_race_39 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_hispanic=case_when(
          race_hispanic == 1 ~ 1,
          race_hispanic == 0 ~ 0,
          is.na(race_hispanic) ~ 0 
          ))

```

##### Asian composite

Composed of individuals who identified as East Asian (e.g., Chinese, Japanese, Korean, Taiwanese), Southeast Asian (e.g., Filipino, Vietnamese, Indonesian), South Asian (e.g., Pakistani, Indian, Nepalese, Sri Lankan), Other Asian; or demograph_race_11, demograph_race_12, demograph_race_13, and demograph_race_19.


```{r Asian composite dummy code}
study1 <- study1 %>%
        mutate(
        race_asian=case_when(
                demograph_race_11 | demograph_race_12 | demograph_race_13 | demograph_race_19 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_asian=case_when(
          race_asian == 1 ~ 1,
          race_asian == 0 ~ 0,
          is.na(race_asian) ~ 0 
          ))

```

##### Native composite

Composed of individuals who identified as Native Hawaiian or Other Pacific Islander, American Indian or Alaskan Native; or demograph_race_41 and demograph_race_42.


```{r Native composite dummy code}
study1 <- study1 %>%
        mutate(
        race_native=case_when(
                demograph_race_41 | demograph_race_42 ~ 1,
                race_multiracial == 1 ~ 0
                ))

study1 <- study1 %>%
        mutate(
          race_native=case_when(
          race_native == 1 ~ 1,
          race_native == 0 ~ 0,
          is.na(race_native) ~ 0 
          ))

```

With these composites in hand, I can look more closely at how many we have in each category

```{r freq race, collapse=TRUE}
table(study1$race_white)
table(study1$race_black)
table(study1$race_asian)
table(study1$race_hispanic)
table(study1$race_native)
table(study1$race_multiracial)
```

There are 152 white participants, 27 Asian, 16 Hispanic, 8 black, 3 native, and no multiracial participants, meaning I can leave multiracial out of the code in Portfolio 10.

#### Quiz recoding

We need to create dummy coded scores for the quiz, so I can get an average more easily. There's also one more question than in Study 2.

```{r quiz question dummy codes}
#Question 1
study1 <- study1 %>%
        mutate(
        Quiz_Q1=case_when(
                Question_1 == 1 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q1=case_when(
          Quiz_Q1 == 1 ~ 1,
          is.na(Quiz_Q1) ~ 0 
          ))

# Question 2
study1 <- study1 %>%
        mutate(
        Quiz_Q2=case_when(
                Question_2 == 1 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q2=case_when(
          Quiz_Q2 == 1 ~ 1,
          is.na(Quiz_Q2) ~ 0 
          ))

# Question 3
study1 <- study1 %>%
        mutate(
        Quiz_Q3=case_when(
                Question_3 == 3 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q3=case_when(
          Quiz_Q3 == 1 ~ 1,
          is.na(Quiz_Q3) ~ 0 
          ))

# Question 4
study1 <- study1 %>%
        mutate(
        Quiz_Q4=case_when(
                Question_4 == 2 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q4=case_when(
          Quiz_Q4 == 1 ~ 1,
          is.na(Quiz_Q4) ~ 0 
          ))

# Question 5
study1 <- study1 %>%
        mutate(
        Quiz_Q5=case_when(
                Question_5 == 3 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q5=case_when(
          Quiz_Q5 == 1 ~ 1,
          is.na(Quiz_Q5) ~ 0 
          ))

#Question 6
study1 <- study1 %>%
        mutate(
        Quiz_Q6=case_when(
                Question_6 == 1 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q6=case_when(
          Quiz_Q6 == 1 ~ 1,
          is.na(Quiz_Q6) ~ 0 
          ))

# Question 7
study1 <- study1 %>%
        mutate(
        Quiz_Q7=case_when(
                Question_7 == 1 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q7=case_when(
          Quiz_Q7 == 1 ~ 1,
          is.na(Quiz_Q7) ~ 0 
          ))
#Question 8
study1 <- study1 %>%
        mutate(
        Quiz_Q8=case_when(
                Question_8 == 2 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q8=case_when(
          Quiz_Q8 == 1 ~ 1,
          is.na(Quiz_Q8) ~ 0 
          ))

#Question 9
study1 <- study1 %>%
        mutate(
        Quiz_Q9=case_when(
                Question_9 == 3 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q9=case_when(
          Quiz_Q9 == 1 ~ 1,
          is.na(Quiz_Q9) ~ 0 
          ))

#Question 10
study1 <- study1 %>%
        mutate(
        Quiz_Q10=case_when(
                Question_10 == 2 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q10=case_when(
          Quiz_Q10 == 1 ~ 1,
          is.na(Quiz_Q10) ~ 0 
          ))

#Question 11
study1 <- study1 %>%
        mutate(
        Quiz_Q11=case_when(
                Question_10 == 1 ~ 1
                ))

study1 <- study1 %>%
        mutate(
          Quiz_Q11=case_when(
          Quiz_Q11 == 1 ~ 1,
          is.na(Quiz_Q11) ~ 0 
          ))
```


Now that I have all the dummy coded versions, I can more easily compute the average score.

```{r quiz average}
study1 <- study1 %>%
        rowwise() %>%
        mutate(
                quiz_average=mean(c_across(Quiz_Q1:Quiz_Q11
        )))

study1$quiz_average <- round(study1$quiz_average, digits = 2)

```



#### Manipulation check

We had three manipulation check items.


##### Disclosure

The first manipulation check item asked about how much the participants believed the questions asked them to disclose, consisting of two items: "Manip_intimacy" and "manip_thoughtful".

First, we should check the reliability.

```{r question discl manip reliability}
alpha(study1[,c("Manip_intimacy", "manip_thoughtful")])

```

Alpha is .7

Now, we can create the composite.

```{r question discl composite}

study1$manip_check_comp2 <- rowMeans(study1[,c("Manip_intimacy", "manip_thoughtful")], 
                              na.rm=TRUE)
```


##### Self-disclosure

The next manipulation check item asked about participants' actual levels of disclosure, consisting of three items: group_tell, group_intimate, group_honest, and group_know. These are terrible names compared to the ones we used in my thesis.

Checking the reliability.

```{r self disclosure manip reliability}
alpha(study1[,c("group_tell", "group_intimate", "group_honest", "group_know")])

```

Reliability is .76

Creating the composite.

```{r self disclosure manip composite}

study1$self_disclosure_comp4 <- rowMeans(study1[,c("group_tell", "group_intimate", "group_honest", "group_know")], 
                              na.rm=TRUE)
```


##### Group disclosure

The final manipulation check item asked about participants' beliefs about their group members' level of disclosure, consisting of three similar items to self-disclosure: group_other_tell, group_other_share, group_other_honest, and group_other_know.

Checking the reliability.

```{r group disclosure manip reliability}
alpha(study1[,c("group_other_tell", "group_other_share", "group_other_honest", "group_other_know")])

```

Reliability is .7

Creating the composite.

```{r group disclosure manip composite}

study1$group_disclosure_comp4 <- rowMeans(study1[,c("group_other_tell", "group_other_share", "group_other_honest", "group_other_know")], 
                              na.rm=TRUE)
```



### Variables of interest


#### Institutional belonging

There are four variables that make up institutional belonging, "school_belong", "school_outsider", "school_comfort", and "school_fit_in".

One of the four variables needs to be reverse scored, school_outsider.

```{r reverse scoring WFU outsider}
study1 <- study1 %>%
        mutate(school_outsiderR= 8- school_outsider
                              )
```

Testing the reliability of the measure. Technically alpha gets masked by ggplot but I can just enable psych for a moment to override it.

```{r institutional belonging reliability}
alpha(study1[,c("school_belong", "school_outsiderR", "school_comfort", "school_fit_in")])

```

Reliability is .89

Creating the actual composite.

```{r institutional belonging composite}

study1$school_belong_comp4 <- rowMeans(study1[,c("school_belong", "school_outsiderR", "school_comfort", "school_fit_in")], 
                              na.rm=TRUE)
```



#### Class belonging

Same as institutional belonging, there are the same four questions (with the prefix class_ this time) and I need to reverse score "class_outsider"

```{r reverse scoring class outsider}
study1 <- study1 %>%
        mutate(class_outsiderR= 8- class_outsider
        )

```

```{r class belonging reliability}
alpha(study1[,c("class_belong", "class_outsiderR", "class_comfort", "class_fit_in")])
```

Reliability is .85

```{r Class belonging composite}
study1$class_belong_comp4 <- rowMeans(study1[,c("class_belong", "class_outsiderR", "class_comfort", "class_fit_in")], 
                              na.rm=TRUE)
```



#### Subject Area motivation

We asked about subject area motivation in two ways: intrinsic motivation and utility motivation.


##### Intrinsic motivation

The intrinsic motivation construct is made of two items: moti_intris_interest and moti_intris_like.

```{r intrinsic motivation reliability}
alpha(study1[,c("moti_intris_interest", "moti_intris_like")])
```

Reliability is .82

```{r intrinsic motivation composite}
study1$motiv_intrins_comp2 <- rowMeans(study1[,c("moti_intris_interest", "moti_intris_like")], 
                              na.rm=TRUE)
```



##### Utility motivation

The utility motivation construct is made of three items: moti_util_now, moti_util_future, and moti_util_grad

```{r utility motivation reliability}
alpha(study1[,c("moti_util_now", "moti_util_future", "moti_util_grad")])
```

Reliability is .91

```{r utility motivation composite}
study1$motiv_util_comp3 <- rowMeans(study1[,c("moti_util_now", "moti_util_future", "moti_util_grad")], 
                              na.rm=TRUE)
```



### Saving data

```{r saving data}
write.csv(study1, file = "Fall_2021_data_R_cleaned.csv")
```




## Finishing up