---
title: "Portfolio Fall 2021 Project data cleaning"
author: "Adam Paul"
date: "3/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> The purpose of this portfolio is to have one final attempt at data cleaning within R. Specifically, I'll work with the data from my project we ran in Fall 2021.

## Set-up

The main packages I need are lavaan and lmertest, but semPlot is used to create a visual model.



library('psych')
library('reshape2')
library('plyr')
library('lavaan')
library('Rcpp')
library('lme4')
library('lmerTest')
library('OpenMx')
library('semPlot')
rm(list=ls())

##### Bringing in the data

> Formerly, I did this with a .sav file,as the data was cleaned using SPSS. This really disrupted the knitting for the markdown, so I tested converting to .csv. The below code was used then.
library(haven)
Classroom_Experiences <- read_sav("~/Master's Program Wake Forest/Research/Online Belonging/FYP materials/Classroom >Experiences Data 8-15-21.sav")
View(Classroom_Experiences_Data_8_15_21)

```{r bringing in data}
Classroom_Experiences <- read.csv("~/Master's Program Wake Forest/Research/Online Belonging/FYP materials/MAP data -csv test.csv")
View(Classroom_Experiences)
```

##### Data cleaning!



