---
title: "Portfolio 4 MAP test- Thesis study 1"
author: "Adam Paul"
date: "3/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> The purpose of this portfolio is to practice with data I already have and prepare for writing a path analysis, so that when I get my data I can run the analyses right away. With that in mind, this markdown has details and information for my own learning, so I can refer to it as I am writing the final version. Which, for some reason, is an earlier portfolio project. Go past me.

## Set-up

The main packages I need are lavaan and lmertest, but semPlot is used to create a visual model.

> Getting semPlot to work was a genuine ordeal, but it should all work now.


library('psych')
library('reshape2')
library('plyr')
library('lavaan')
library('Rcpp')
library('lme4')
library('lmerTest')
library('OpenMx')
library('semPlot')
rm(list=ls())

##### Bringing in the data

> Formerly, I did this with a .sav file,as the data was cleaned using SPSS. This really disrupted the knitting for the markdown, so I tested converting to .csv. The below code was used then.
library(haven)
Classroom_Experiences <- read_sav("~/Master's Program Wake Forest/Research/Online Belonging/FYP materials/Classroom >Experiences Data 8-15-21.sav")
View(Classroom_Experiences_Data_8_15_21)

```{r bringing in data}
Classroom_Experiences <- read.csv("~/Master's Program Wake Forest/Research/Online Belonging/FYP materials/MAP data -csv test.csv")
View(Classroom_Experiences)
```

##### Data cleaning!

For the purposes of Portfolio 4, I am not cleaning the data. However, in the full thesis, I will clean the data here.

The focus of this portfolio is ensuring that I understand how to run a path analysis and put together the code for the model, so I am using data I already cleaned.

------------------

# MEDIATION MODELS

------------------

> This section is information given to me by Shannon, kept here and in other path analysis portfolios so I have access to them.

Reminder: Rules of thumb to evaluate goodness of fit: CFI/TLI > 0.95, RMSEA < 0.05, SRMR < 0.06  (Reference: Hu & Bentler, 1999)

### References: 
> http://davidakenny.net/cm/fit.htm
 https://www.psychologie.uzh.ch/dam/jcr:ffffffff-b371-2797-0000-00000fda8f29/chisquare_diff_en.pdf
 Chi-Square Distribution Table: http://sites.stat.psu.edu/~mga/401/tables/Chi-square-table.pdf
 http://www.structuralequations.com/resources/Basic_lavaan_Syntax_Guide_Aug1_2013.pdf

------------------

# Current settings 

------------------

### Resampling 
 Right now, bootstraps are set at 1000 resamples. This to make running the script not take impossibly long
 When running final models, 10000 resamples would probably be preferred - but it'll take a while. 

### Type of CIs
 lavaan can handle four types of bootstrapped CIs; add "boot.ci.type = " to parameterEstimates
 options for types are "norm", "basic", "perc" (percentile), and "bca.simple" 
 defualt is "perc" (I think )

### Significance Levels
 alpha level for bootstrapping can be adjusted; add "level = " to parameterEstimates 
 options for levels are e.g., "95", "99"

------------------

## Model time!

------------------

##### Hypotheses

H1: Students who disclose in a class or small group will report higher feelings of belonging in the class and at their institution.

H2: Prompting disclosure in class or in small groups will be associated with greater disclosure and greater sense of belonging at the classroom and institutional level, as well as whether the relationship between disclosure and institutional belonging is mediated by classroom belonging.

>These are the covariates I am including in the model:
Gender, with female as the reference class- 
        gender_identity_dich 
Race, with white as the reference class- 
        race_asian race_black race_hispanic race_multiracial 
Classmates and group members known prior
        classmates_know groupdesc_know_prior

#### Creating the model!


```{r model creation}
model_1 <-   '  #Class Belonging direct effects
                belong_class_comp4 ~ a*cdiscl_disclose + gender_identity_dich + race_asian 
                + race_black + race_hispanic + race_multiracial + classmates_know +                          groupdesc_know_prior
                
                belong_class_comp4 ~ b*gdiscl_disclose
                
                #WFU Belonging direct effects
                belong_wfu_comp4 ~ c*cdiscl_disclose + gender_identity_dich + race_asian 
                + race_black + race_hispanic + race_multiracial + classmates_know +                          groupdesc_know_prior
                
                belong_wfu_comp4 ~ d*gdiscl_disclose
                
                #Mediator for class disclosure
                cdiscl_disclose ~ e*cdiscl_prompt_discl
                belong_class_comp4 ~ cdiscl_prompt_discl
                belong_wfu_comp4 ~ cdiscl_prompt_discl
                
                #Mediator for group disclosure
                gdiscl_disclose ~ f*gdiscl_do_emot
                belong_class_comp4 ~ gdiscl_do_emot
                belong_wfu_comp4 ~ gdiscl_do_emot
                
                #removing correlation between prompted disclosures
                cdiscl_prompt_discl ~~ 0*gdiscl_do_emot 
                '
```

#### Testing the fit


```{r testing fit}
fit <- sem(model_1, data=Classroom_Experiences)

summary(fit, fit.measures=TRUE, standardize=TRUE, rsquare=TRUE)
```

#### Graphing!

> This graph is kinda hideous, but it's what I've got for now. I don't like the model outputting in the smaller form, so I will usually pull it out of the chunk. 

```{r Path visualization}
semPaths(fit, whatLabels="par", layout="tree", nCharNodes=0, sizeMan=12, sizeMan2=6)
```


>Currently unused code, because I don't have the time to swap the order. It was fine, and then the order swapped.
nodeLabels= c("Class Belonging", "WFU Belonging", "Class Disclosure", "Group Disclosure", "Group Prompted Disc", "Class Prompted Disc", "Group Members Known", "Classmates Known", "Multiracial", "Hispanic", "Black", "Asian", "Gender")

#### Bootstrapping

Bootstrapping is currently set to 1000 for the sake of time, not because it's what I will ultimately want to run. With the real data, I will go for 10000.

#####for boot strapped 95% CIs

```{r bootstrapped code}
fit <- sem(model_1, data=Classroom_Experiences, se="bootstrap", bootstrap = 1000)
summary(fit, fit.measures=TRUE, standardize=TRUE, rsquare=TRUE)
```

# What does the data mean?

>This section will be used for me to write up what I've learned from the data analysis, as a point to reference the findings while I'm working later.

>H1: Students who disclose in a class or small group will report higher feelings of belonging in the class and at their institution.

Group disclosure was a significant predictor of belonging at both levels, but class disclosure was only a significantly negative predictor of institutional belonging!

>H2: Prompting disclosure in class or in small groups will be associated with greater disclosure and greater sense of belonging at the classroom and institutional level, as well as whether the relationship between disclosure and institutional belonging is mediated by classroom belonging.

Classroom prompting
        Significantly predicted actual disclosure, but was not a significant predictor of class or institutional belonging.
Group Prompting
        Significantly predicted actual disclosure, but was not a significant predictor of class or institutional belonging.
